Generating the Blueprints of the Java Ecosystem

### 2015-02-27 msr.tex
one pass

Software metrics provide a means to extract useful and measurable information about the structure of a software system. Software development artifacts consist of digital data, many of their aspects can be easily measured. This is exemplified by the early appearance of metrics like {\sc loc} (Lines of Code) and {\sc cloc} (Comment Lines of Code). Over the years, a wealth of different metrics have appeared in the literature. Many of them, such as Chidamber and Kemerer's~\cite{CHKE94} and the metrics presented by Lorenz et al.~\cite{LOKI94} can be automatically calculated by static analysis tools.

In this paper, we present a dataset that contains popular metrics, calculated from the analysis of a large collection of Java projects. All projects were taken from the {\it Maven Central Repository}~\cite{MAVEN}. Maven is a build automation tool used mostly for Java projects. To build a software component, it dynamically downloads Java libraries and Maven plug-ins from the Maven central repository, and stores them in a local cache. The repository can be updated with new projects and also with new versions of existing projects that can depend on other versions. Note that this repository has been previously analyzed by researchers to produce relevant datasets as we discuss in Section~\ref{sec:rel}. To analyze the projects coming from this repository, we used the following tools: a) {\sc ckjm} \cite{Spi05g}, b) {\sc jd}epend~\cite{JDEPEND}, and c) {\sc clmt}~\cite{SGKL09}. All tools focus on three main aspects of a software system, namely: object-oriented design, program size, and package design.

@@

Maven is a build automation tool used primarily for Java projects, and it is maintained by the Apache Software Foundation~\cite{MAVEN}. To describe the software project being built, its dependencies on other libraries, and the build order Maven uses {\sc xml}. The central repository contains more than 400,000 {\sc jar}s, in a variety of languages that are all using the {\sc jvm} platform as their runtime environment. These languages include Java, Clojure, Groovy, and Scala.

@@

The dataset construction process is illustrated in Figure~\ref{fig:dataset-construction}. Initially, a snapshot of the Maven repository was downloaded locally. The repository contains various projects and their releases. We identified only the Java projects. The repository contains several versions of each project. All versions were filtered out and only the latest was kept. Each project consists of two {\sc jar} files, one that contains a compiled version of the project and one with the source code. The projects that did not had both binary and source jars, were excluded from the experiment.

When the selection process was finished, the {\sc jars} were downloaded from the central repository and with the help of a queue, delivered into a series of workers, which executed the tools and provided a series of measurements stored in {\sc json} form. The result files were validated by scripts and then the data were imported in a MySQL database.

Table \ref{tbl:oss-size-metrics} presents several size metrics regarding the final version corpus. It included 11,365 projects, with more than 110 million lines of code. Overall the dataset contains almost 33 million unique measurements.

Three tools were selected to calculate the measurements.

@@

Table~\ref{tbl:selected-metrics} presents the key metrics that are calculated and stored in the dataset. Each tool focuses on a different aspect of a software system; \textit{ckjm} focuses on object-oriented design metrics, \textit{jdepend} on package design, while \textit{clmt} focuses on program size metrics. Several metrics are calculated by more than one tool, like \textit{cyclomatic complexity}, which is calculated by both \textit{clmt} and \textit{ckjm}. Both calculations are available in the database.

A series of secondary metrics are also stored in the database. For example, a set of metrics that count the usage of specific {\sc dsl} application libraries in a software project. These are not included in the Table~\ref{tbl:selected-metrics}, since they are not commonly accepted. In Section~\ref{sec:dsl}, we describe a small experiment based on these metrics that measures the popularity of {\sc dsl} usage for the dataset anf therefore for the Java ecosystem in general.

@@

The measurements provided by this dataset, can be used both by researchers and practictioners to validate their ideas. Researchers can use the data to experiment with new quality models, or just validate them in real projects. Practitioners who develop several tools that calculate metrics that already exist in this dataset, can use it to validate that their tools work correctly.

@@

@@

This dataset followed a narrow selection process, identifying and analysing only projects that were written in Java and had the source code and the binary jar available in the repository. The latter reduced significantly the number of projects that were analysed, since many of them provided only the binary jar. Since only \textit{clmt} analyses the source code, this ended in drastically reduction of measurements that were exposing object-oriented and package design issues.

@@

In reference \cite{MKLGS14}, Mitropoulos et al. also experimented in a similar way, by running the FindBugs~\cite{HP04} tool on a large part of the maven repository. Our work differs from these two approaches since it presents the metrics calculated by the three aforementioned tools; \textit{ckjm}, \textit{jdepend}, and \textit{clmt}.

>>>>

Software metrics provide a means to extract useful and measurable information about the structure of a software system. Software development artifacts consist of digital data, thus many of their aspects can be easily measured. This is exemplified by the early appearance of metrics like {\sc loc} (Lines of Code) and {\sc cloc} (Comment Lines of Code). Over the years, a wealth of different metrics have appeared in the literature. Many of them, such as Chidamber and Kemerer's~\cite{CHKE94} and the metrics presented by Lorenz et al.~\cite{LOKI94} can be automatically calculated by static analysis tools.

In this paper, we present a dataset that contains popular metrics, calculated from the analysis of a large collection of software artifacts written in Java. All artifacts were taken from the {\it Maven Central Repository}~\cite{MAVEN}. Maven is a build automation tool used mostly for Java projects. To build a software component, it dynamically downloads Java libraries and Maven plug-ins from the Maven central repository, and stores them in a local cache. The repository can be updated with new projects and also with new versions of existing projects that can depend on other versions. To analyze the projects coming from this repository, we used the following tools: a) {\sc ckjm} \cite{Spi05g}, b) {\sc jd}epend~\cite{JDEPEND}, and c) {\sc clmt}~\cite{SGKL09}. All tools focus on three main aspects of a software system, namely: object-oriented design, program size, and package design.

@@

Maven is a build automation tool used primarily for Java projects, and it is maintained by the Apache Software Foundation~\cite{MAVEN}. To describe the software project being built, its dependencies, and the build order, Maven uses {\sc xml}. The central repository contains more than 400,000 {\sc jar}s, in a variety of programming languages such as Java, Clojure, Groovy, and Scala. All languages use the {\sc jvm} platform as their runtime environment.

@@

The dataset construction process is illustrated in Figure~\ref{fig:dataset-construction}. Initially, a snapshot of the Maven repository was downloaded locally. The repository contains various projects that contain several versions. In our experiment, we only used Java projects. For every Java project all versions were filtered out and only the latest was kept.

Every project consists of two {\sc jar} files, one that contains a compiled version of the project and one that contains the source code. The projects that did not had both binary and source jars, were excluded from the experiment.

When the selection process was finished, we created a series of processing tasks based on the containing all the selected {\sc jar}s, and added them to a task queue mechanism.  Then we executed a number of workers
% dimitro@bkarak: written in?
that checked out tasks from the queue, applied the static analysis tools on each {\sc jar} and stored the results to the data repository (a {\sc m}y{\sc sql} database system). The tools that were applied on every {\sc jar} were the following.

@@

% When the selection process was finished, the {\sc jar}s were downloaded from the central repository and with the help of a queue, delivered into a series of workers, which executed the tools and provided a series of measurements stored in {\sc json} form. The result files were validated by scripts and then the data were imported in a MySQL database.

Table \ref{tbl:oss-size-metrics} presents several size metrics regarding the final corpus we used. It included 11,365 projects, with more than 110 million lines of code. Overall the dataset contains almost 33 million unique measurements.

Table~\ref{tbl:selected-metrics} presents the key metrics that were calculated and stored in the dataset. Each tool focuses on a different aspect of a software system; \textit{ckjm} focuses on object-oriented design metrics, \textit{jdepend} on package design, while \textit{clmt} focuses on program size metrics. Several metrics are calculated by more than one tool, like \textit{cyclomatic complexity}, which is calculated by both \textit{clmt} and \textit{ckjm}. Both calculations are available in the database.

% A series of secondary metrics wesre also stored in the database. For example, a set of metrics that count the usage of specific {\sc dsl} application libraries in a software project. These are not included in the Table~\ref{tbl:selected-metrics}, since they are not commonly accepted. dimitro@bkarak: then why do you mention it?

In Section~\ref{sec:dsl}, we describe a small experiment based on these metrics that measures the popularity of {\sc dsl} usage for the dataset anf therefore for the Java ecosystem in general.

@@

The measurements provided by this dataset, can be used both by researchers and practictioners to validate their ideas. Researchers can use the data to experiment with new quality models, or just validate them in real projects. Practitioners who develop several tools that calculate metrics that already exist in this dataset, can use it to validate if their tools work correctly.

@@

% dimitro@bkarak: yes what?

@@

This dataset followed a narrow selection process, identifying and analysing only projects that were written in Java and had the source code and the binary {\sc jar} available in the repository. The latter reduced significantly the number of projects that were analysed, since many of them provided only the binary {\sc jar}. Since only \textit{clmt} analyses the source code, this ended in drastically reduction of measurements that were exposing object-oriented and package design issues.

@@

In reference \cite{MKLGS14}, Mitropoulos et al. also experimented in a similar way, by running the FindBugs~\cite{HP04} tool on a large part of the maven repository. Our work differs from these two approaches since it presents the metrics calculated by the three different static analysis tools.
